SPANG: Template Library for Making SPARQL Queries FAIR

### 2020-05-23 paper.tex
v1.8

\title{SPANG Template Library: Making SPARQL Queries FAIR}
\titlerunning{SPANG: Making SPARQL Queries FAIR}
\author{Hirokazu Chiba}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
How can we maximize the value of accumulated SPARQL queries? 
%SPARQL is a key component of the Semantic Web, thus the reusability of SPARQL is important.
An increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows interoperable framework for querying the RDF datasets; however, reuse of SPARQL queries remains limited due to the lack of accessibility and findability of SPARQL queries written in various projects. 
%SPARQL contributes to the standardized interface to the RDF, and maximizing the reusability of RDF; however, reusability of SPARQL itself remains to be addressed. 
In this study, we focus on FAIR (findable, accessible, interoperable,
reusable) principle to improve reusability of SPARQL queries.
We present the process of making SPARQL queries FAIR and demonstrate the usefulness of the resulting query library.
%FAIRification of the SPARQL is a plausible way to make SPARQL interoperable and reusable.
Here, we define a SPANG library as a controleld query set with specific
extension of SPARQL so that the query set follows FAIR pinciple.
%Here we propose a SPARQL extension to construct a reusable query library to enrich the Semantic Web platform.
Our contribution is showing steps for making SPARQL queries FAIR;
1) assigning a metadata set for annotating a SPARQL query, which
enables to identify SPARQL query and provides useful information for the query;
2) parameterize SPARQL queries, which makes each query reusable in different settings;
3) assigning globally unique and stable identifier for each query;
4) construct a server to provide the index of SPARQL queries.
This study make SPARQL queries FAIR, and leads to the construction of eco-system where SPARQL queries are reused to maximize the potential of the Semantic Web.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

First, we review FAIR principles and describe how to apply the principle to SPARQL.
Further, we demonstrate how to use the SPANG templete library.
Then, we describe the details of SPANG template specifications.

>>>>

\title{SPANG: Template Library for Making SPARQL Queries FAIR}
\titlerunning{SPANG: Template Library for Making SPARQL Queries FAIR}
\author{Hirokazu Chiba}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
How can we maximize the value of accumulated SPARQL queries? 
%SPARQL is a key component of the Semantic Web, thus the reusability of SPARQL is important.
An increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows interoperable framework for querying the RDF datasets; however, reuse of SPARQL queries remains limited due to the lack of accessibility and findability of SPARQL queries written in various projects. 
%SPARQL contributes to the standardized interface to the RDF, and maximizing the reusability of RDF; however, reusability of SPARQL itself remains to be addressed. 
In this study, we focus on FAIR (findable, accessible, interoperable,
reusable) principle to improve reusability of SPARQL queries.
We present the process of making SPARQL queries FAIR and demonstrate the usefulness of the resulting query library.
%FAIRification of the SPARQL is a plausible way to make SPARQL interoperable and reusable.
Here, we define a SPANG library as a controleld query set with specific
extension of SPARQL so that the query set follows FAIR pinciple.
%Here we propose a SPARQL extension to construct a reusable query library to enrich the Semantic Web platform.
Our contribution is showing steps for making SPARQL queries FAIR;
1) assigning a metadata set for annotating a SPARQL query, which
enables to identify SPARQL query and provides useful information for the query;
2) parameterize SPARQL queries, which makes each query reusable in different settings;
3) assigning globally unique and stable identifier for each query;
4) construct a server to provide the index of SPARQL queries.
This study make SPARQL queries FAIR, and leads to the construction of eco-system where SPARQL queries are reused to maximize the potential of the Semantic Web.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{SPANG Overview}
%%%%%%%%%%%%%%%%%%%%%%%%

In this study, we review FAIR principles and describe how to apply the principle to SPARQL.
Further, we demonstrate how to use the SPANG templete library.
Then, we describe the details of SPANG template specifications.

@@
